My title

### 2024-02-19 paper/paper.qmd
Committing changes to paper.qmd and paper.pdf Feb 19 @ 1:47pm

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction
In recent years the United States has experienced a shift in racial framing and valence within media and society. Despite similar rates of illicit drug use among Black and White Americans throughout history, Black Americans are significantly more likely to be incarcerated for drug offenses. This phenomenon is particularly pronounced in the context of the war on drugs @BoboJohnson2004. In contrast, White Americans, especially in the context of the opioid crisis, experience a more lenient, medicalized approach to their drug use, revealing a racialized American drug policy @NetherlandHansen2017.

This shift in the racial framing of drug use and policy within media and societal discourse is a direct result of White Americans and the the opioid epidemic. As a result, drug abuse has been met with a more sympathetic, medicalized response, highlighting a stark contrast in how drug use has historically been conceptualized and addressed throughout the country. A 2022 paper by Tanika Raychaudhuri, Tali Mendelberg, and Anne McDonough in The Journal of Politics
Volume 85, Number 1 (January 2023), titled "The Political Effects of Opioid Addiction Frames" delves into this interplay of media coverage on public support for drug treatment policies and the racial framing and valence (positive or negative portrayal) of opioid abuse influence on policy views @Raychaudhuri2022.

Raychaudhuri, Mendelberg, and McDonough's study investigates the impact of media coverage on public support for drug treatment policies, focusing on how the racial framing and valence of opioid abuse influence White Americans' policy views. Unlike previous drug epidemics portrayed negatively and associated with non-White users, the opioid crisis is often framed sympathetically and as predominantly affecting White Americans. This sympathetic portrayal is thought to contribute to the broad public support for treating opioid abuse as a public health issue rather than a criminal one @Raychaudhuri2022.

The study revealed that when opioid users, both White and Black, were shown in a sympathetic light, White people's support for treatment options rose. This supportive attitude was noticeably stronger when the users depicted were White, pointing to a bias in empathy based on race. On the other hand, portrayals lacking sympathy didn't really change opinions, regardless of the user's race @Raychaudhuri2022.

@@

This reproduction was conducted using the statistical programming language R @citeR. To further enable the analysis, the following packages were employed: knitr @knitr for dynamic report generation and kableExtra @kableExtra to enhance the presentation of tables with advanced formatting options. The broom package @broom for converting statistical analysis objects into tidy formats, aiding in more effective and efficient data handling and interpretation. dplyr @dplyr For data manipulation and preparation. ggplot2 @ggplot2 for creating effective and well presented data visualizations, enabling proper communication of findings. Lastly, tidyverse @Wickham2019.

# Data
### Source
The paper used for replication is from the The Journal of Politics Volume 85, Number 1
(January 2023) which examines whether sympathy for drug users is racially selective among White Americans @Raychaudhuri2022. This Reproduction seeks to utilize one of the studies major findings and then asses whether ideological tendencies influences the response to sympathetic media frames about opioid abuse, potentially conditioning the effect of racial selectivity in sympathy. This could help understand if conservative and liberal individuals are equally affected by the racial framing of opioid abuse stories.

### Literature
Much of the studies literature review draws on Michelle Alexander's 2010 book that focuses on mass incarceration in the age of colourblindness @Alexander2010. The paper also draw upon Michael Collins 2019 Article Titled “No Gentler War on Drugs: How Race Made the Opioid Crisis" @Collins2019. These peices of literature aid in theorizing that historically, drug abuse focuses on the racial impact of the “War on Drugs,” a punitive policy response to crack cocaine. Unlike the opioid crisis, the crack cocaine crisis led to the incarceration of millions, most of whom were non-White Americans @Raychaudhuri2022. Media coverage of cocaine often depicted Black users in an unsympathetic light, linking their drug use to involvement in other crimes. Julie Netherland and Helena Hansen's article titled "White opioids: Pharmaceutical race and the war on drugs that wasn’t" aids in proving that by contrast to the crack epidemic, news coverage now frames opioid abuse largely as a crisis afflicting White Americans and when doing so it also tends to use a positive valence. By contrast, the Black “face” often carries negative overtones @Raychaudhuri2022. Overall, the literature aids in building the study's foundation argument that the racialization of drug epidemics can not merely be looked at as a matter of public health crises but also as an issue deeply intertwined with socio-political narratives and racial prejudices.

### Methodology
This paper will replicate the survey data that was originally collected for the (2022) paper by Raychaudhuri, Mendelberg, and McDonough. The study constructed a hypothetical newspaper story about opioid abuse, varying the race of the users (Black or White) and the valence of the frame (sympathetic or unsympathetic). The focus on Black and White radicalized groups was guided by the White “face” of the opioid epidemic in contrast to the Black “face” of other drug epidemics @Raychaudhuri2022. The authors then conducted two randomized between-subjects survey experiments on two separate samples. The main sample consists of 1,517 White American adults from the NORC Amerispeak Panel, a national probability based sample. The other sample was from the survey firm Dynata. The authors describe the process as follows: "Each participant was randomly assigned to one of four treatments or a no-story control. In the treatments, respondents read a hypothetical news story about drug use, varying the race of users (White or Black) and the valence of the frame (sympathetic or unsympathetic)" @Raychaudhuri2022.

# Features
The original survey data assessed participants on 45 variables, being representative of geographical location, education level, polictical affiliation, ideology, and racial attitudes and affiliations. The main sample consists of 1,517 White American adults from the NORC Amerispeak Panel, a national probability based sample. The other sample was from the survey firm Dynata. The original paper includes three data sets; data from pre-test (Dynata panel), data from main experiment (NORC Amerispeak Panel), and data for the state-level opioid mortality analysis (CDC). However, this reproduction is focused solely on the data from the main experiement and includes variables only found in the main experiment data set. This reproduction omits data for the state-level opioid mortality analysis (CDC) and the data from the pre-test (Dynata panel). This data sets were not relevant to the reproduction. This reproduction also changed the name of some of the main experiments variables in the data set for better clarity. The data set was then cleaned to include only relevant variables applicable to this reproduction. @tbl-cleaned_data shows a sample of the cleaned data set.

@@

The terms `conservative:media_frames` and `liberal:media_frames` are designed to detect whether the relationship between media framing and policy support differs across the ideological spectrum. The response variable `policy` is binary, indicating whether an individual supports or rejects treatment-focused intervention for opioid abuse. Therefore, a logistic regression model (logit model) is employed to predict the probability of policy support, which is suitable for binary outcome data. The logit model is defined as:

@@

The next step after setting up the logistic regression model is understanding how people's ideological beliefs and the various ways the media frames opioid use impact their support for policy . To do this, a coefficient table (@tbl-coefficients) was produced to capture the essence of the relationships between predictor variables. The goal is to quantify the relationship between each predictor variable (ideological beliefs, media framings, etc.) and interaction term (like conservative:media_frames) and the corresponding likelihood of policy support.

The coefficients, for instance, point to whether a factor, like political leaning or a specific type of media narrative, nudges public opinion towards or away from a certain policy stance, and importantly, how strong that nudge is. It also looks at standard errors, z-values, and p-values to check the confidence in these effects—are they solid or could they be just chance findings? And those 95% confidence intervals give a ballpark figure, showing where the true effect likely falls. @tbl-terms_explanation_table summarizes each term found in the coefficient table. Furthermore, the model's fit statistics (@tbl-fit-statistics), including AIC and BIC among others, offer a snapshot of the model's overall performance and suitability for the data at hand.

@@

```{r}
#| warning: false
#| echo: false
#| label: tbl-coefficients
#| tbl-cap: Logistic regression coefficient estimates, standard errors, z-values, p-values, and 95% confidence intervals.

@@

```{r}
#| warning: false
#| echo: false
#| label: tbl-fit-statistics
#| tbl-cap: Model fit statistics including Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC), and other relevant measures.

@@

# Creating a bar graph using ggplot
ggplot(data_long_melted, aes(x = category, y = proportion, fill = policy_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("support_for_treatment" = "blue", "support_for_arrest" = "red")) +
  labs(title = "Policy Support by Valence and Race (Media Frames)",
       x = "Valence and Race",
       y = "Proportion",
       fill = "Policy Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

# Create the Forest Plot
ggplot(tidy_model, aes(x = estimate, y = term)) +
  geom_point() + # Add points for estimates
  geom_errorbarh(aes(xmin = estimate - std.error, xmax = estimate + std.error), height = 0.2) + # Error bars
  theme_minimal() + # Use a minimal theme
  labs(x = "Estimate", y = "Variable", title = "Ideology Model Estimates") + # Labels
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") # Vertical line at zero
```

# Discussion
The interest to examine the interaction effects between political ideology (conservative or liberal) and media framing (sympathetic or unsympathetic, towards white or black opioid users) on the support for policy regarding opioid abuse treatment was heavily influenced by the reading of Pyra M, Taylor B, Flanagan E, Hotton A, Johnson O, Lamuda P, Schneider J, and Pollack H's study titled "Support for evidence-informed opioid policies and interventions: The role of racial attitudes, political affiliation, and opioid stigma" @Pyra2022. The significance of the study was that it emphasized that despite the opioid epidemic being portrayed as less racially and politically divisive compared to previous drug crises, underlying racial and political attitudes remain influential @Pyra2022. Their study revealed that Democratic Party affiliation (political ideological outlook of "liberal" in the sense of this paper) correlated with greater support for these measures and mandatory treatment, possibly as an alternative to more punitive approaches.

@@

### Findings
The null hypothesis ($H_0$) stated that there would be no interaction effect between political ideology and media framing on support for opioid abuse treatment policy. However, @fig-interactions shows varying estimates across different combinations of political ideology and media framing. This suggests that there may be an interaction effect. This is because the estimates for both liberal and conservative ideologies when paired with certain sympathetic framings are positive and do not overlap with the zero line, indicating a significant effect. This suggests that there is indeed some interaction effect, which leads to the rejection of the null hypothesis.

@@

### Accounting for Bias
Accounting for bias is a critical component of research arise naturally when collecting quantitative and qualitative data. Raychaudhuri, Mendelberg, and McDonough took important steps to mitigate bias in their experiments by pre-registering their methods and employing a fully controlled design with realistic news stories. This approach helps prevent confirmation bias, meaning that the authors were not influenced by their own beliefs and values to unintentionally back their hypotheses because of the way their studies were conducted. While the authors did present measures to control for bias, there are still potential areas where bias could be present. This could include sampling bias as the authors make no mention to whether the study captures a diverse cross-section of White Americans in terms of factors like socioeconomic status, education level, or demographics.

Other accounts for bias could be present in the media content used for the study. The authors state that the generated news stories used in the study were constructed based on realistic, pre-existing new stories. Therefore, If the realistic news stories used in the study are based on actual content that already has a bias, then the study could inadvertently measure the effects of existing media bias rather than pure racial or valence framing effects. Furthermore, the wording of news stories used in the experiment could influence respondents' answers. Subtle language cues can lead to different interpretations, affecting the study's outcomes. In fact, the entire piece of media is left to interpretation.

@@

### Future Research 
Future research could look into the psychological mechanisms that may explain why and how political ideologies modulate responses to media framing of opioid users. It is also imperative to understand the role of individual identity, such as race and personal experience with opioid abuse, in shaping one's reaction to policy framing. Could it be that personal connections to the opioid crisis, either through community impact or personal acquaintance with sufferers, are powerful enough to override political predispositions? Personal experience could play a significant role in an individuals views on media framing and policy and seems like a strong future area of research regarding both this paper and the original.

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "∗Code and data are available at: https://github.com/Bellamaclean7/Ideological-Influences-of-Opioid-Addiction-Media-Frames; Replication on Social Science Reproduction platform available at: https://doi.org/10.48152/ssrp-5x1j-c733"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction
In recent years the United States has experienced a shift in racial framing and valence within media and society concerning drug epidemics. Despite similar rates of illicit drug use among Black and White Americans throughout history, Black Americans are significantly more likely to be incarcerated for drug offenses. This phenomenon is particularly pronounced in the context of the war on drugs [@BoboJohnson2004]. In contrast, White Americans, especially in the context of the opioid crisis, experience a more lenient, medicalized approach to their drug use, revealing a racialized American drug policy [@NetherlandHansen2017].

This shift in the racial framing of drug use and policy within media and societal discourse is a direct result of White Americans and the the opioid epidemic. As a result, drug abuse has been met with a more sympathetic, medicalized response, highlighting a stark contrast in how drug use has historically been conceptualized and addressed throughout the country. A 2022 paper by Tanika Raychaudhuri, Tali Mendelberg, and Anne McDonough in The Journal of Politics
Volume 85, Number 1 (January 2023), titled "The Political Effects of Opioid Addiction Frames" delves into the impact of media coverage on public support for drug treatment policies, focusing on how the racial framing and valence of opioid abuse influence White Americans' policy views. Unlike previous drug epidemics portrayed negatively and associated with non-White users, the opioid crisis is often framed sympathetically and as predominantly affecting White Americans. This sympathetic portrayal is thought to contribute to the broad public support for treating opioid abuse as a public health issue rather than a criminal one [@Raychaudhuri2022].

The study revealed that when opioid users, both White and Black, were shown in a sympathetic light, White people's support for treatment options rose. This supportive attitude was noticeably stronger when the users depicted were White, pointing to a bias in empathy based on race. On the other hand, portrayals lacking sympathy didn't really change opinions, regardless of the user's race [@Raychaudhuri2022].

@@

This reproduction was conducted using the statistical programming language R [@citeR]. To further enable the analysis, the following packages were employed: knitr [@knitr] for dynamic report generation and kableExtra [@kableExtra] to enhance the presentation of tables with advanced formatting options. The broom package [@broom] for converting statistical analysis objects into tidy formats, aiding in more effective and efficient data handling and interpretation. dplyr [@dplyr] For data manipulation and preparation. ggplot2 [@ggplot2] for creating effective and well presented data visualizations, enabling proper communication of findings. Lastly, tidyverse [@Wickham2019].

# Data
### Source
The paper used for replication is from the The Journal of Politics Volume 85, Number 1
(January 2023) which examines whether sympathy for drug users is racially selective among White Americans [@Raychaudhuri2022]. This Reproduction seeks to utilize one of the studies major findings and then asses whether ideological tendencies influences the response to sympathetic media frames about opioid abuse, potentially conditioning the effect of racial selectivity in sympathy. This could help understand if conservative and liberal individuals are equally affected by the racial framing of opioid abuse stories.

### Literature
Much of the studies literature review draws on Michelle Alexander's 2010 book that focuses on mass incarceration in the age of colourblindness [@Alexander2010]. The paper also draw upon Michael Collins 2019 Article Titled “No Gentler War on Drugs: How Race Made the Opioid Crisis" [@Collins2019]. These peices of literature aid in theorizing that historically, drug abuse focuses on the racial impact of the “War on Drugs,” a punitive policy response to crack cocaine. Unlike the opioid crisis, the crack cocaine crisis led to the incarceration of millions, most of whom were non-White Americans [@Raychaudhuri2022]. Media coverage of cocaine often depicted Black users in an unsympathetic light, linking their drug use to involvement in other crimes. Julie Netherland and Helena Hansen's article titled "White opioids: Pharmaceutical race and the war on drugs that wasn’t" aids in proving that by contrast to the crack epidemic, news coverage now frames opioid abuse largely as a crisis affecting White Americans and when doing so it also tends to use a positive valence. By contrast, the Black “face” often carries negative overtones [@Raychaudhuri2022]. Overall, the literature aids in building the study's foundation argument that the racialization of drug epidemics can not merely be looked at as a matter of public health crises but also as an issue deeply intertwined with socio-political narratives and racial prejudices.

### Methodology
This paper will replicate the survey data that was originally collected for the (2022) paper by Raychaudhuri, Mendelberg, and McDonough. The study constructed a hypothetical newspaper story about opioid abuse, varying the race of the users (Black or White) and the valence of the frame (sympathetic or unsympathetic). The focus on Black and White radicalized groups was guided by the White “face” of the opioid epidemic in contrast to the Black “face” of other drug epidemics [@Raychaudhuri2022]. The authors then conducted two randomized between-subjects survey experiments on two separate samples. The main sample consists of 1,517 White American adults from the NORC Amerispeak Panel, a national probability based sample. The other sample was from the survey firm Dynata. The authors describe the process as follows: "Each participant was randomly assigned to one of four treatments or a no-story control. In the treatments, respondents read a hypothetical news story about drug use, varying the race of users (White or Black) and the valence of the frame (sympathetic or unsympathetic)" [@Raychaudhuri2022].

# Features
The original survey data assessed participants on 45 variables, being representative of geographical location, education level, polictical affiliation, ideology, and racial attitudes and affiliations. The main sample consists of 1,517 White American adults from the NORC Amerispeak Panel, a national probability based sample. The other sample was from the survey firm Dynata. The original paper includes three data sets; data from pre-test (Dynata panel), data from main experiment (NORC Amerispeak Panel), and data for the state-level opioid mortality analysis (CDC). However, this reproduction is focused solely on the data from the main experiement and includes variables only found in the main experiment data set. This reproduction omits data for the state-level opioid mortality analysis (CDC) and the data from the pre-test (Dynata panel). This reproduction also changed the name of some of the main experiments variables in the data set for better clarity. The data set was then cleaned to include only relevant variables applicable to this reproduction. @tbl-cleaned_data shows a sample of the cleaned data set.

@@

The terms `conservative:media_frames` and `liberal:media_frames` are designed to detect whether the relationship between media framing and policy support differs across the ideological spectrum. The response variable `policy` is binary, indicating whether an individual supports or rejects treatment-focused intervention for opioid abuse. Therefore, a logistic regression model (logit model) is employed to predict the probability of policy support, which is suitable for binary outcome data. The logit model is defined as follows:

@@

The next step after setting up the logistic regression model is understanding how people's ideological beliefs and the various ways the media frames opioid use impact their support for policy. To do this, a coefficient table (@tbl-coefficients) was produced to capture the essence of the relationships between predictor variables. The goal is to quantify the relationship between each predictor variable (ideological beliefs, media framings, etc.) and interaction term (like conservative:media_frames) and the corresponding likelihood of policy support.

The coefficients, for instance, point to whether a factor, like ideological leaning and a specific media narrative, nudges public opinion towards or away from a certain policy stance, and importantly, how strong that nudge is. It also looks at standard errors, z-values, and p-values to check the confidence in these effects—are they solid or could they be just chance findings? Lastly, the table records the 95% confidence intervals, which give a ballpark figure showing where the true effect likely falls. @tbl-terms_explanation_table summarizes each term found in the coefficient table. Furthermore, the model's fit statistics (@tbl-fit-statistics), including AIC and BIC among others, offer a snapshot of the model's overall performance and suitability for the data at hand.

@@

```{r}
#| warning: false
#| echo: false
#| fig.pos: 'H'
#| label: tbl-coefficients
#| tbl-cap: Logistic regression coefficient estimates, standard errors, z-values, p-values, and 95% confidence intervals.

@@

```{r}
#| warning: false
#| echo: false
#| fig.pos: 'H'
#| label: tbl-fit-statistics
#| tbl-cap: Model fit statistics including Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC), and other relevant measures.

@@

# Creating a bar graph using ggplot
ggplot(data_long_melted, aes(x = category, y = proportion, fill = policy_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("support_for_treatment" = "blue", "support_for_arrest" = "red")) +
  labs(title = "Policy Support by Valence and Race (Media Frames)",
       x = "Valence and Race",
       y = "Proportion",
       fill = "Policy Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

# Create the Forest Plot
ggplot(tidy_model, aes(x = estimate, y = term)) +
  geom_point() + # Add points for estimates
  geom_errorbarh(aes(xmin = estimate - std.error, xmax = estimate + std.error), height = 0.2) + # Error bars
  theme_minimal() + # Use a minimal theme
  labs(x = "Estimate", y = "Variable", title = "Ideology Model Estimates") + # Labels
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") # Vertical line at zero

# Discussion
The interest to examine the interaction effects between political ideology (conservative or liberal) and media framing (sympathetic or unsympathetic, towards white or black opioid users) on the support for policy regarding opioid abuse treatment was heavily influenced by the reading of Pyra M, Taylor B, Flanagan E, Hotton A, Johnson O, Lamuda P, Schneider J, and Pollack H's study titled "Support for evidence-informed opioid policies and interventions: The role of racial attitudes, political affiliation, and opioid stigma" [@Pyra2022]. The significance of the study was that it emphasized that despite the opioid epidemic being portrayed as less racially and politically divisive compared to previous drug crises, underlying racial and political attitudes remain influential [@Pyra2022]. Their study revealed that Democratic Party affiliation (political ideological outlook of "liberal" in the sense of this paper) correlated with greater support for these measures and mandatory treatment, possibly as an alternative to more punitive approaches.

@@

### Findings
The null hypothesis ($H_0$) stated that there would be no interaction effect between political ideology and media framing on support for opioid abuse treatment policy. However, @fig-interactions shows varying estimates across different combinations of political ideology and media framing. Estimates for both liberal and conservative ideologies when paired with certain sympathetic framings are positive and do not overlap with the zero line, indicating a significant effect. This suggests that there is indeed some interaction effect, which leads to the rejection of the null hypothesis.

@@

The data reveals that political ideology and media framing can influence attitudes towards treatment-focused policies. This is especially true when liberals are exposed to media that frames a white opioid user in a sympathetic lens, providing partial support for the first hypothesis $H_1$. The most interesting find was conservatives positive relationship regarding support for treatment related policy when faced with media framing black opioid users in an unsympathetic lens. In summary, this analysis reveals that political ideology does somewhat interact with media framing to influence support for treatment related policies, the interactions require further research and analysis to draw stronger conclusions. It is important to note that the lack of statistical significance for most comparisons, due to confidence intervals crossing the zero point, indicates a need for caution in over-interpreting these results.

### Accounting for Bias
Accounting for bias is a critical component of research and arises naturally when collecting quantitative and qualitative data. Raychaudhuri, Mendelberg, and McDonough took important steps to mitigate bias in their experiments by pre-registering their methods and employing a fully controlled design with realistic news stories. This approach helps prevent confirmation bias, meaning that the authors were not influenced by their own beliefs and values to unintentionally back their hypotheses because of the way their studies were conducted. While the authors did present measures to control for bias, there are still potential areas where bias could be present. This could include sampling bias as the authors make no mention to whether the study captures a diverse cross-section of White Americans in terms of factors like socioeconomic status, education level, or demographics.

Other accounts for bias could be present in the media content used for the study. The authors state that the generated news stories used in the study were constructed based on realistic, pre-existing new stories. Therefore, if the realistic news stories used in the study are based on actual content that already has a bias, then the study could inadvertently measure the effects of existing media bias rather than pure racial or valence framing effects. Furthermore, the wording of news stories used in the experiment could influence respondents' answers. Subtle language cues can lead to different interpretations, affecting the study's outcomes. In fact, the entire piece of media is left to interpretation.

@@

### Future Research 
Future research could look into the psychological mechanisms that may explain why and how political ideologies modulate responses to media framing of opioid users. It is also imperative to understand the role of individual identity, such as personal experience with opioid abuse, in shaping one's reaction to policy. Could it be that personal connections to the opioid crisis, either through community impact or personal acquaintance with sufferers, are powerful enough to override political predispositions? Personal experience could play a significant role in an individuals views on media framing and policy and seems like a strong future area of research regarding both this paper and the original.
